\section{Scenario Customization}
\subsection{Configuration File Syntax}

All parameters are integers unless specified otherwise.

\subsection{Ship Parameters}
\begin{itemize}
\item[turns]  Ship maneuverability (turn radius)
\item[torp-damage]  Torpedo damage at zero radius
\item[torp-blast-range]  Torpedo blast radius - range at which damage = 0
\item[mine-damage] Mine damage at zero radius
\item[mine-blast-range]  Mine blast radius - range at which damage = 0
\item[phaser-damage] Phaser damage at zero radius
\item[phaser-range] Phaser range - range at which damage = 0
\item[ph-pulses]  Number of time slices that a phaser will lock on a target
\item[torp-speed]  Torpedo speed
\item[max-speed] Maximum ship speed
\item[shield-repair]  Shield repair rate
\item[hull-repair]  Damage repair rate
\item[max-fuel]  Maximum fuel capacity
\item[torp-cost] Torpedo fuel cost (unused) .  Note that weapons costs are
calculated from the damage caused by each weapon
\item[mine-cost] Mine fuel cost (unused)
\item[phaser-cost] Phaser fuel cost (unused)
\item[det-cost] Detonate other torps fuel cost
\item[warp-cost] Warp engine fuel cost
\item[cloak-cost] Cloak fuel cost
\item[recharge]  Fuel recharge rate
\item[accint]  Acceleration rate
\item[decint]  Deceleration rate
\item[max-armies]  Maximum army capacity
\item[weapon-cool]  Weapon cooling rate
\item[engine-cool]  Engine cooling rate
\item[max-damage] Maximum structural damage sustainable before destruction
\item[shield-max]  Maximum shield damage absorption
\item[teleport-heat]  Teleport engine heat cost
\item[teleport-cost]  Teleport fuel cost
\item[teleport-range]  Teleport radius
\item[turbo-speed]  Turbo warp speed (unused)
\item[turbo-time]  Turbo warp duration (unused)
\item[reload] time between torp salvos
\item[burst] number of shots in a torp salvo
\item[phaser-fail]  Probability of phaser failure.  Similarly for 
{\tt torp-fail}, {\tt trans-fail}, {\tt shield-fail}, 
 {\tt cloak-fail}, {\tt lrsensor-fail}, {\tt srsensor-fail}, {\tt lock-fail},
and {\tt cooling-fail}.
\end{description}

\subsection{Robot Parameters}
\begin{description}
\item[hscruise] speed robots go when going to assist another robot.
\item[cruise] speed robots patrol at.
\item[battle] speed robots go in battle.
\item[flee] speed robots run away at.
\item[cloaked] speed robots go when cloaked.
\item[refresh] speed robots go when they want to cool down.
\item[engage] distance at which a robot starts firing at a target.
\item[disengage] distance at which a robot stops firing at a target.
\item[shotdamage] damage a robot's salvo does.
\item[circledist] distance at which a robot would like to stand off it's
target.
\item[sneaky] chance that, on creation, a robot will elect to make cloaked
attacks.
\item[max-cloak] distance at which a robot, making a sneaky attack, will
cloak.
\item[min-cloak] distance at which a robot, making a sneaky attack, will uncloak.
\end{description}

\subsection{Empire Parameters}
\begin{description}
\item[empire] String. Empire's name.
\item[icon] String parameter specifies bitmap file which is in the defined LIBDIR.  Bitmap is 
automatically rotated through all 16 orientations.  The bitmap should be specified
with front pointing upwards, and some slight aliasing will result.
\item[robotname]  String parameter specifies robot name.
\end{description}

\subsection{Planet Parameters}
\begin{description}
\item[planet] String. Planet's name.
\item[empire] String. Planet's empire.
\item[home]  Boolean flag.  Specifies that planet is the home planet of the
empire.
\item[fuel] Boolean flag.  Specifies that planet is a fuel source; fuel is
replenished at twice normal planetary rates when planet is orbited. 
\item[repair] Boolean flag.  Specifies that planet is a repair yard; damage is
repaired at twice normal planetary rates when planet is orbited. 
\item[x]  Specifies X coordinate of planet.  Similarly for {\tt y}.
\item[armies] Specifies number of armies on planet at start of game.
Note that a large number of  armies makes the planet an effective obstacle;
any ship passing through such a planet will be destroyed.
\end{description}

\subsection{Global Parameters}
\begin{description}
\item[robot-giveup-time] If after this many seconds there are no players
in the game the robots quit. Defaults to 10.
\item[death-time] Player is dead for this number of seconds.  Defaults to 4. 
\item[torp-life-min]   Torpedoes live for at least this number of seconds.  
Defaults to 7.  Similarly for {\tt mine-life-min} (default is 60).
\item[torp-life-var]  Torpedoes may live for at least this number of seconds 
in addition to the minimum lifetime.  Defaults to 2. Similarly for 
{\tt mine-life-var} (default is 120).
\item[player-explode-time] Player explodes for this number of seconds.
\item[weapon-lock-min] Overheated weapons are unusable for at least this
number of seconds  up   to an additional {\tt weapon-lock-var} seconds.
Defaults are 10 and 15. Similarly for {\tt engine-lock-min} and 
{\tt engine-lock-var}.  Defaults are  identical to those of weapons.
\item[self-destruct-time]  Self-destruct countdown time.
\item[robot-giveup-time] Robots will give up within this number of
seconds if no combat is initiated.
\item[orbit-speed]   Maximum warp allowed while entering orbit. 
\item[det-dist]  Effective range of the counter-torpedo batteries. 
\item[orbit-dist]  Range at which a ship can enter orbit. Defaults to 900.
\item[planet-fire-dist]  Range at which a planet can damage a ship.
Defaults to 1500.
\item[phaser-hit-angle]  Number of degrees that a phaser will hit.
\item[phaser-hit-langle]  Number of degrees that a locked phaser will hit.
\item[auto-quit]  Auto-logout countdown time.
\item[cool-penalty]   Floating point number.  Specifies the increase
in weapon and engine cooling rate when both shields and cloak are inactive.
Defaults to 1.25.
\item[fast-destruct]  Boolean flag.  If set, self-destruct time does not
increase with alert level.   Default is to double self-destruct time for
yellow alert, and double again for red alert.
\item[mine-detonate]  Boolean flag.   If set, mines detonate when
destroyed.  This feature is potentially very nasty.  It also allows instant
self-destruction, cheating players out of kills.
\end{description}
