\thispagestyle{empty}
\ 
\hbox{\colastitledisp
\vbox{
\vspace{3cm}
\begin{center}
\fboxrule 0.4pt \fboxsep 1pt
\fbox{\fboxrule 3pt \fboxsep 30pt \fbox{\Huge\bf GWM Manual}}
\end{center}
\vspace{2cm}
\begin{center}
\huge
The X11 {\bf G}eneric {\bf W}indow {\bf M}anager
\end{center}
\vspace{2cm}
\begin{center} 
\Large
$Revision: 1.5 $\\
%$Date: 89/11/28 11:35:02 $
\end{center}
\vspace{2cm}
\begin{center}
\LARGE\sf Colas Nahaboo\\
\large\tt colas@mirsa.inria.fr
\end{center}
\vspace{1cm}
%\begin{center}
%\LARGE\sf Vania Joloboff\\
%\large\tt vania@mirsa.inria.fr
%\end{center}
\vspace{1cm}
\begin{center}
\copyright BULL 1989
\end{center}
}}

\newpage

\section*{Copyright restrictions}
{\bf\begin{flushleft} 
Copyright 1989 GROUPE BULL\\
Copyright 1989 Massachusetts Institute of Technology
\end{flushleft}}

{\sf
Permission to use, copy, modify, and distribute this software and its
documentation for any purpose and without fee is hereby granted, provided
that the above copyright notice appear in all copies and that both that
copyright notice and this permission notice appear in supporting
documentation, and that the name of GROUPE BULL not be used in advertising
or publicity pertaining to distribution of the software without specific,
written prior permission.  GROUPE BULL makes no representations about the
suitability of this software for any purpose.  It is provided ``as is''
without express or implied warranty.

GROUPE BULL disclaims all warranties with regard to this software,
including all implied warranties of merchantability and fitness,
in no event shall GROUPE BULL be liable for any special,
indirect or consequential damages or any damages
whatsoever resulting from loss of use, data or profits,
whether in an action of contract, negligence or other tortious
action, arising out of or in connection with the use 
or performance of this software.
}

\section*{Acknowledgements}

Although I am the only designer and implementor of {\GWM}, the project could
not have been completed without the help of many individuals, namely:

Vania Joloboff, who created the KOALA BULL team at INRIA, and started the 
{\GWM} project, and KOALA members Vincent Bouthors and Daniel Dardailler
who tested {\GWM} and wrote {\WOOL} profiles.

Gilles Kahn, Francis Montagnac, and the CROAP team at INRIA who intensively
used {\GWM} and provided me with machines to develop it. A special thanks
goes to Janet Incerpi who corrected the documentation.

All my beta testers: Raphael Bernhard, Didier Poirot, Matthieu Devin,
Christophe Muller, Laurent Hill, Anne Marie Bustos, Simon Kaplan ...
and last, but certainly not least, Bob Scheifler himself!

And since July 1989, {\GWM} has been enhanced by the growing community of
the {\tt gwm-talk} mailing list, whose members are alas too numerous to be all 
cited here, but are giving me an invaluable feedback.

\begin{flushright}
{\Large Colas Nahaboo.\quad}
{\small
KOALA Project -- BULL Research c/o INRIA\\
2004 route des Lucioles -- 06565 Valbonne Cedex -- FRANCE\\
}
\end{flushright}

\section*{Support}

\sloppy You can mail any questions relative to {\GWM} by electronic mail to
{\tt gwm@mirsa.inria.fr}\footnote{Where you can mail also requests to
be added on the two mailing lists: {\tt gwm-talk}, for general comments, and
{\tt gwm-bugs}, for reporting (and be warned of reported) bugs},
and you can find the latest {\GWM} version by anonymous ftp on expo.lcs.mit.edu
(18.30.0.212) and on trinidad.inria.fr (192.5.60.125) where the mailing lists
are also archived.
